---
title: "Reproducible Research: Peer Assessment 1"
output: 
html_document:
  keep_md: true
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```
## **Loading and preprocessing the data**
### Load and look at the data format and structure.

```{r load and look at data}
data <- read.table("activity 2.csv", header = TRUE, sep = ",")
head(data)
str(data)
```
### Convert date column from factor to date.

```{r Date Conversion}
data$date <- as.Date(data$date, format = "%m/%d/%y")
str(data)
```
## **What is the mean total number of steps taken per day?**
### 1. Calculate the total number of steps taken per day.
```{r}
library(dplyr)
sum_data <- data %>% group_by(date) %>% 
  summarize(total = sum(steps, na.rm = TRUE))
head(sum_data)
```
### 2. Make a histogram of the total number of steps taken each day. 

```{r histogram}
hist(sum_data$total, xlab = "Total Steps", main = "Total Steps Taken Per Day")
```


### 3. Calculate and report the mean and median of the total number of steps
### taken per day. 

```{r mean and median}
summ_data <- sum_data %>%
  summarize(mean = mean(total, na.rm = TRUE), median = median(total, na.rm = TRUE))
head(summ_data)
```

## **What is the Average Daily Activity Pattern**
### 1. Make a time series plot of the intervals and average steps taken
```{r}
data$interval <- as.factor(data$interval)
str(data)
library(ggplot2)
TS_data <- data %>% group_by(interval) %>% summarize (mean = mean(steps, na.rm = TRUE))
head(TS_data)
ggplot(TS_data, aes(as.numeric(interval), mean)) + 
  geom_point() + 
  geom_path() +
  xlab("5 Minute Intervals Beginning at 0 = 12AM") +
  ylab("Average Steps") +
  ggtitle("Average Steps Taken Per Day in 5 Minute Intervals")
```


### 2. Which interval, on average across all days, contains the maximum number
### of steps?
```{r}
max <- TS_data[which.max(TS_data$mean),]
max
```
## **Imputing Missing Values**

### 1. Calculate and report the total number of NAs in the dataset.
```{r Total NAs}
colSums(is.na(data))
```
### 2.and 3.Replace NA values with the mean value across days during that interval and create a new data set including missing values filled in

```{r Replace Missing Values}
data_NA <- data %>% 
  group_by(interval) %>%
  mutate(steps_fill = ifelse(is.na(steps),
                             mean(steps, na.rm = TRUE), steps))

head(data_NA)
```
### 4a. Histogram showing total number of steps per day with avarage missing values
```{r Total Steps}
sumNA_data <- data_NA %>% group_by(date) %>% 
  summarize(total = sum(steps_fill, na.rm = TRUE))
head(sumNA_data)
hist(sumNA_data$total, xlab = "Total Steps", main = "Total Step Per Day with Replaced Missing Values")
```

### 4b.Mean and median steps per day
```{r}
summNA_data <- sumNA_data %>%
  summarize(mean = mean(total, na.rm = TRUE), median = median(total, na.rm = TRUE))
head(summNA_data)
```
## ** Are there differences in activity patterns between weekdays and weekends?**
### 1. Create a new factor variable in the dataseet with two levels
```{r}
data_NA$date <- as.Date(data_NA$date, format = "%m/%d/%y")
str(data)
weekdays1 <- c('Monday', 'Tuesday', 'Wednesday', 'Thursday','Friday')
data_NA$days <- factor((weekdays(data_NA$date) %in% weekdays1), 
         levels=c(FALSE, TRUE), labels=c('weekend', 'weekday'))
head(data_NA)
```
### 2. Make a panel plot. 

```{r}
summ_days <- data_NA %>% 
  group_by(interval, days) %>% 
  summarize(Average = mean(steps_fill))
summ_days
```

```{r}
ggplot(summ_days, aes(as.numeric(interval), Average)) + 
  geom_point() + geom_path() +
  facet_grid(days ~ .) + xlab("Time in 5-minute Intervals (starting at 0 = 12AM") +
  ylab("Average Steps")
```
